---
title: Tutorial 03
subtitle: "ENVX1002"
description: The University of Sydney
date: today
date-format: "[Semester 1,] YYYY"
---

# R Commands for Binomial and Poisson
$X \sim Bin(np)$ and  $X \sim Po(\lambda)$

| Function              | Binomial        | Poisson            |
|:--------------------- |:--------------- |:------------------ |
| PDF $P(X=x)$          | $dbinom(x,n,p)$ | $dpois(x,\lambda)$ |
| CDF $F(x)=P(X\le{x})$ | $pbinom(x,n,p)$ | $ppois(x,\lambda)$ |


## Binomial distribution

1. Let's think about a situation that could be modelled by $X \sim Bin(n = 8, p = 0.7)$?

2. Use R to find the following **Binomial** probabilities for hybrid tomatoes:

The new Dixie Red hybrid tomato is claimed to have larger fruit, better disease package and higher yield potential compared to the current industry leader Florida 47 R hybrid tomato.

<https://www.growingproduce.com/vegetables/seminis-introduces-dixie-red-fresh-tomato-variety/>

Suppose we plant 8 seeds of Dixie Red and that each seed has a 70% success rate of germination.

- Let $X =$ The number of Dixie Red seeds that germinate. What is the distribution of $X$?
- Use R to find the exact probability distribution function and add in the table.

| Number seeds $x$     | 0   | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8
|:-------------------- |:--- |:--- |:--- |:--- |:--- |:--- |:--- |:--- |:--- 
| Probability $P(X=x)$ |     |     |     |     |     |     |     |     |

```{r, eval = TRUE}
x <- c(0,1,2,3,4,5,6,7,8) # creates x values. You can also use x <- 0:8
p <- dbinom(x,8,0.7) #calculated the probabilities at each x.
rbind(x,round(p,3)) # view the probabilities with x in top row and rounded p in bottom row

sum(p) # Notice that all the probabilities sum to 1
```

- Use R to construct a histogram of the pdf. You can hit the code button to reveal the code to use.

```{r, eval = TRUE}
plot(x,p,type="h") # type "h" draws lines to the points
```

Notice the Shape: The pdf is **left skewed**. This is because there is a high success rate *(p = 0.7)* so we expect more germination than no germination.

- What is the probability that 7 or more Dixie Red seeds germinate?

$P(X\ge7)=P(X=7)+P(X=8)=1-P(X<7)=1-P(X\le6)$

$P(X\ge7)=\frac{8!}{7!(8 - 7)!}0.7^7(1-0.7)^{8-7} + \frac{8!}{8!(8 - 8)!}0.7^8(1-0.7)^{8-8}$ 

```{r}
dbinom(7,8,0.7)+dbinom(8,8,0.7)

## or

1-pbinom(6,8,0.7)
```


## Poisson distribution

On a certain joy flight path, the number of dolphin pod sightings is 2.2 per hour. Using a Poisson model, calculate the probabilities of viewing 0 to 8 dolphin pods in the table below

- Use R to findnd the exact probability distribution function and all in the table.

| Number dolphin pods $x$ | 0   | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   
|:----------------------- |:--- |:--- |:--- |:--- |:--- |:--- |:--- |:--- |:---
| Probability $P(X=x)$    |     |     |     |     |     |     |     |     |    

```{r}
x <- c(0,1,2,3,4,5,6,7,8) # creates x values. You can also use x <- 0:8
p <- dpois(x,2.2) #calculated the probabilities at each x and round to 3 d.p.
rbind(x,round(p,3)) # view the probabilities with x in top row and p in bottom row

sum(p) ## notice that all the probabilities don't sum to 1
```

- do a barplot of the probabilities

```{r, eval = TRUE}
plot(x,p,type="h") # type "h" draws lines to the points
```

- What is the probability of seeing no dolphin pods.

$P(X=0)=\frac{2.2^0e^{-2.2}}{0!}=0.1108$

so about 11% of the flights i.e. 1 in 10 flights you will see no dolphins

```{r}
dpois(0,2.2)
```

- What is the probability of seeing more than 3 dolphin pods?

$P(X\ge3)=1-P(X=0)-P(X=1)-P(X=2)-P(X=3)$

$P(X\ge3)=1-\frac{2.2^0e^{-2.2}}{0!}-\frac{2.2^1e^{-2.2}}{1!}-\frac{2.2^2e^{-2.2}}{2!}-\frac{2.2^3e^{-2.2}}{3!}$
$=0.1806476$

```{r}
1-dpois(0,2.2)-dpois(1,2.2)-dpois(2,2.2)-dpois(3,2.2)

1-ppois(3,2.2)
```
